---
title: "hw2ExploreTESDataSet"
author: "Haoyu(Michael)_Wang"
date: '2022-09-14'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Discription of the name of the columns
`age` represent the age of the employee,
`gender` represent the gender of the employee,1	Male	2	Female
`satisfaction` represent the employee's satisfaction of the job,
`jobCharacteristicsVTM` represent the most important characteristics of the job,
`YoW` represent the years altogether you've worked for your present employer
`likenessOfPromotionInF` represent the employee's feeling of are they to be promoted'
`decisionAffectYourOwn` represent the involvedness of the employee making decision affecting their own job,
`paticipationOfBudgetaryDecision` represent do the employee participate in budgetary decisions?
`proudness` shows the employees proud of their job or not,
`lessMoneyToStay` represent how likely the employee willing to turn down another higher salary job to stay in the organization,
`relationshipBtwColleague` represent the personal feeling of the between the colleagues.

## Exploring Start
#### Reading in csv file.
```{r}
myTesDF<-read.csv("C:\\Users\\whylo\\Desktop\\MITA_Third_Year\\Business Forecasting\\Homeworks\\hw2\\Typical_Employee_Survey_Data.csv")
```
#### Showing the information of the data  
```{r}
names(myTesDF)
ncol(myTesDF)
nrow(myTesDF)

dim(myTesDF)
head(myTesDF)
```
#### Brief of the DF before fatorization
```{r}
str(myTesDF)
```
#### Brief of the DF after fatorization
```{r}

myTesDF$gender<-as.factor(myTesDF$gender)
myTesDF$satisfaction<-as.factor(myTesDF$satisfaction)
myTesDF$jobCharacteristicsVTM<-as.factor(myTesDF$jobCharacteristicsVTM)
myTesDF$likenessOfPromotionInF<-as.factor(myTesDF$likenessOfPromotionInF)
myTesDF$decisionAffectYourOwn<-as.factor(myTesDF$decisionAffectYourOwn)
myTesDF$paticipationOfBudgetaryDecision<-as.factor(myTesDF$paticipationOfBudgetaryDecision)
myTesDF$proudness<-as.factor(myTesDF$proudness)
myTesDF$lessMoneyToStay<-as.factor(myTesDF$lessMoneyToStay)
myTesDF$relationshipBtwColleague<-as.factor(myTesDF$relationshipBtwColleague)

str(myTesDF)
```
#### Showing summary of the DF
```{r}
summary(myTesDF)
```
#### Displaying part of the DF by columns and rows
```{r}
#columns 1 to 3
myTesDF[1:3]
#rows No.4
myTesDF[4,]
#combined
myTesDF[2:11,c("age","gender","YoW","proudness")]
myTesDF[3:8,3:5]
```
#### Levels
```{r}
table(myTesDF$jobCharacteristicsVTM)
```
#### Filtering 
```{r}
#multiple  filtering
tesDF_subesetOfMaleandJCVTMInRange<-subset(myTesDF,gender == 1 & (jobCharacteristicsVTM==3 | jobCharacteristicsVTM==4))
tesDF_subesetOfMaleandJCVTMInRange
```
#### Ordering 
```{r}
#rev provides a reversed version of its argument. It is generic function with a default method for vectors and one for dendrograms.
#This providing the employee with the longest years of working for this org with the least satisfaction.
partofDF<- myTesDF[order(myTesDF$YoW,rev(myTesDF$satisfaction),decreasing = TRUE),]
partofDF[1:24,1:5]

```
#### Chage name of the columns
```{r}
#satisfaction->happiness
names(myTesDF)[3] <- "hapiness"
head(myTesDF,3)
#happiness->satisfaction
names(myTesDF)[3] <- "satisfaction"
head(myTesDF,3)
```
## Changing age data for row 33
```{r}
myTesDF[33,1]
#was 40
myTesDF[33,1] <- 22
myTesDF$age[33]
#now 22
myTesDF$age[33] <- 40
myTesDF[33,1]
#changing back to 40
```
#### Graphing
```{r}

library(magrittr)
library(dplyr)
library(ggplot2)
ggpStackedBarchart<-myTesDF %>% group_by(satisfaction,jobCharacteristicsVTM) %>% count(satisfaction) %>%
  ggplot(aes(x = satisfaction, y = n,fill=jobCharacteristicsVTM))+geom_text(aes(label=n)) +geom_bar(stat = "identity")




ggpStackedBarchart<-ggpStackedBarchart+labs(y="Count",title="stacked barchart of nubmer of different job characteristicsVTM group by satisfaction")
ggpStackedBarchart

```

```{r}
myTesDF2<-myTesDF
myTesDF2["ageGroup"] = cut(myTesDF2$age, c(20, 31, 42, 53, 64), c("20-31", "32-42", "43-53","53-64"), include.lowest=TRUE)
str(myTesDF2)
```

```{r}
myTesDF2 %>% group_by(ageGroup) %>% count(ageGroup)%>%
  ggplot( aes(x="", y=n, fill=ageGroup)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)+theme_void()+labs(title="pie chart of age group")

```

```{r}
stem(myTesDF$age,scale = .5)
hist(myTesDF$YoW)
boxplot(myTesDF$age)
```


#### Summary stats
```{r}
min(myTesDF$YoW)
max(myTesDF$YoW)
range(myTesDF$YoW)
rangeOfDataFrame = max(myTesDF$YoW)-min(myTesDF$YoW)
rangeOfDataFrame
mean(myTesDF$age)
sd(myTesDF$age)
var(myTesDF$age)
sqrt(myTesDF$age)
fivenum(myTesDF$age)
IQR(myTesDF$age)
quantile(myTesDF$age)
summary(myTesDF$age)
## 
## n	
## the number of non-NA observations in the sample.
## 
## conf	
## the lower and upper extremes of the ‘notch’ (if(do.conf)). See the details.

## out	
## the values of any data points which lie beyond the extremes of the whiskers (if(do.out)).
##
## Note that $stats and $conf are sorted
##
boxplot.stats(myTesDF$age)
boxplot.stats(myTesDF$age)$conf


# Data Frame Summary
summary(myTesDF)
by(myTesDF$age,myTesDF$satisfaction,mean)
by(myTesDF$age,myTesDF$jobCharacteristicsVTM,sd)
by(myTesDF$YoW,myTesDF$likenessOfPromotionInF,summary)
by(myTesDF$YoW,myTesDF$likenessOfPromotionInF,mean)
by(myTesDF$age,myTesDF$proudness,mean)
by(myTesDF$age,myTesDF$likenessOfPromotionInF,mean)

```


```{r}
aggregate(myTesDF$age,list("Type" = myTesDF$satisfaction),mean)
aggregate(myTesDF$YoW,list("Type" = myTesDF$proudness),mean)
```


```{r}

```


```{r}

```


```{r}

```